---
title: "ICE 1"
output: html_notebook
---

# Part 0

If you haven't already, install the latest version of R from
[r-project.org](http://www.r-project.org). Also install the latest
version of RStudio Desktop from [rstudio.com](http://www.rstudio.com).

From ICON, download the zip file associated with ICE 1. Once downloaded,
decompress the zip file onto your computer's drive. Navigate to the
directory which contains the decompressed contents, and then start
RStudio by double clicking on the *ice.Rproj* file. (If you're reading
this file on your computer, you probably have already completed this
step!)

# Part 1

Uncomment and execute the two lines in the next chunk of code.

```{r}



```

# Part 2

Calculate the sum of 10 and 5. This can either be done in the notebook
or in the console. For commands executed within a notebook, it is often
a good idea to wrap the command inside the print(...) command to ensure
that the output is printed when the notebook is sourced.

```{r}



```

Calculate the square root of `pi`. If you need help, try
`help("groupGeneric")` and `?pi` in the console (not in the notebook).

```{r}



```

In the next chunk of code, write the following lines of code to save
`sqrt(pi)` in the variable `x` and to print the value of `x`:

    x <- sqrt(pi)
    print(x)

Press the *Run > Restart R and Run All Chunks* command in RStudio to run
your entire notebook so far. If done correctly, you should see `x` in
the Environment pane and the value of `x` printed in the notebook.

```{r}



```

In this notebook, calculate square of variable `x` value.

```{r}



```

Use the inequality character `<` to determine if the reciprocal of `x`
is less than 0.1.

```{r}



```

In the console, investigate `example(Sys.Date)` and `help("Sys.Date")`
to learn how to use the `Sys.Date(...)` command. Then, in the notebook,
calculate and print the date 173 days ago. Wrap with the `print(...)`
command.

```{r}



```

# Part 3

R includes by default a package called `datasets`, which contains many
interesting data sets. In the console, issue the command `library(help
= "datasets")` to see an entire list of the data sets provided by this
package.

```{r}



```

One of the data sets is called `state.x77`. In this notebook, uncomment
and run the following command, which loads `state.x77` into an object
called `states` (we will learn more about the `as.data.frame(...)`
command in later modules):

```{r}



```

Examine the first few rows of states in either the console or in the
notebook. If you need help, try `help("head")`.

```{r}



```

Examine a statistical summary of states by using the command
`summary(...)`.

```{r}



```

Examine the structure of states using the command `str(...)`.

```{r}



```

Calculate the mean of the population of all 50 states. If you need help,
try `help(mean)`. In order to reference the `Population` column of the
data set, use the `$` operator, i.e., `states$Population`.

```{r}



```

Use the command

    plot(states$Illiteracy, states$`Life Exp`)

to plot life expectancy on the y-axis against illiteracy on the x-axis.
For help, try `?plot`, and note that, since the column for life
expectancy contains a space, we have to use the back-tick marks.

```{r}



```

# Part 4

In the console, update all of your packages via the `update.packages()`
command. Such a command would not typically be put in a notebook.

```{r}



```

Install the `swirl` package in the console. For help, try
`?install.packages` or use the RStudio Packages window pane. Note that
commands to install packages are not typically executed from a notebook.

```{r}



```

In the notebook, make sure the `swirl` package is loaded using the
`library(...)` command. Note: Such commands *are* usually put in
notebooks.

```{r}



```

In the console, type `swirl()`. Then follow the instructions to install
the *R Programming* course, select it, and complete the *Basic Building
Blocks* lesson.

# Part 5

Use RStudio's graphical interface to import `allstate.csv`. You can also
read about the data in the included PDF file *allstate_description.pdf*.

As an alternative method, now use the `library(readr)` command and
`read_csv(...)` to import the data. Save the data frame in a variable
called `df`.

```{r}



```

Use the code `View(df)` from the console to get a spreadsheet-like view
of the data frame.

Examine the structure of `df` using `str(...)`.

```{r}



```

Examine the column names of `df`. Also look at the number of columns.

```{r}



```

Examine the number of rows of `df`.

```{r}



```

Calculate the total memory used by `df` in megabytes.

```{r}



```

# Part 6

Examine the seventh entry of the third column of `df` by first
extracting it as a data frame (using single square brackets) and then as
just the entry itself (using double square brackets).

```{r}



```

Use the `:` symbol to select rows `7` to `11` and columns `3` to `5`
into a smaller data frame called df2 this display it.

```{r}



```

Now use the `c(...)` command (combine function) to select rows `7` and `11` (not "7 to 11")
and columns `state` and `cost` and save this in df3 then display it.

```{r}



```

# Part 7

Prepare a summary of the `car_age` column using the `summary(...)`
command. Use either the `$` notation or double square brackets to
extract the column itself as a vector.

```{r}



```

Prepare a histogram of the `cost` column. See `?hist` for help.

```{r}



```

Use `plot(...)` to plot `cost` on `car_age`. Extract the columns each
using the `$` notation.

```{r}



```

# Part 8

Count the number of `NA` in the `duration_previous` column using
`is.na(...)`.

```{r}



```

Count the total number of entries in `df` that are `NA`.

```{r}



```

Reread the data into `df4` with: (1) column `customer_ID` as `character`;
(2) the following definition of the option `na`, which will ensure that
both `"NA"` and `"0"` are read as `NA`: `c("NA", "0")`

```{r}



```

Now, recount the total number of entries in `df` that are `NA`.

```{r}



```

Also calculate the pecentage of entries in `df` that are `NA`.

```{r}



```

# Part 9

Investigate the data in `boston2013.csv` to answer the following
questions regarding the Boston Marathon in 2013, which was interrupted
by a terrorist bombing. You may have to reload the data by executing the
`read_csv` command again. Save your commands in this notebook.

What was the finish time (in minutes) of the fastest runner?

```{r}



```

At what approximate time (relative to the 24-hour clock) did the bomb
explode?

```{r}



```

# Part 10

Read the SPSS file `spss.sav` into a data frame using the function
`read.spss` in the package `foreign`. The foreign package contains functions that will allow 
you to import data files from some of the most commonly used statistical software packages 
such as SAS, Stata and SPSS. 

```{r}



```
